---
title: | 
  | BST 6200 Spatial Statistics and Disease Mapping
  | Homework 2
author: "Miao Cai <miao.cai@slu.edu>"
date: "3/2/2020"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

1. Consider the bramblecanes data from class. For each of the three ages of bramble canes (0, 1, or 2) construct the following:

    a) a kernel density estimate and display it with a heat map (use trial and error to get an appropriate bin width)
    b) a hexagonal bin plot (SKIP)
    c) an estimate of the K function along with the envelope for testing CSR.

```{r message=FALSE, warning=FALSE}
# a) Kernel density estimate
pacman::p_load(GISTools, tmap, tmaptools, spatstat, sf)

data( bramblecanes )
bramblecanes$marks = as.character(bramblecanes$marks)

p_kde = function(group_num = "0"){
  group0 = as(bramblecanes[bramblecanes$marks == group_num], "SpatialPoints")
  proj4string(group0) = CRS('+init=epsg:26978')
  group0_dens = smooth_map(group0)
  
  p = tm_shape(group0_dens$raster) +
    tm_raster() + 
    tm_layout(title = paste0("Mark: ", group_num))
  return(p)
}

tmap_arrange( p_kde("0") , p_kde("1"), p_kde("2"))
```

```{r}
# c) An estimate of the K function
pacman::p_load(fMultivar)
bramb_env = envelope(bramblecanes, Kest, correction = "border")
plot(bramb_env)
```


2. Consider the location of trees in the trees.csv data set given on Blackboard.

    a) Create a ppp object that has region [0,200] by [0,200].
    b) The tree diameters are given in the variable dia. Make a histogram for the tree diameters.
    c) Construct a kernel density and display it with a heat map.
    d) Display the kernel density with a set of contours.  (SKIP)
    e) Estimate the K function and plot it along with the envelope for testing CSR.
    f) Consider only the trees with diameter less than or equal to 20. Repeat part (e).
    
```{r}
# a) a ppp object
trees = data.table::fread("homework/trees.csv")
trees_ppp = trees[,ppp(x, y, xrange = c(0, 200), yrange = c(0, 200))]
class(trees_ppp)
```

```{r}
# b) tree diameters
pacman::p_load(ggplot2)
trees %>%
  ggplot(aes(x = dia)) +
  geom_histogram(aes(y = ..density..),
    color = "#1F3552",
    fill = "#4271AE",
    size = 0.7) +
  geom_density(size = 1)
```


```{r}
# c) Construct a kernel density and display it with a heat map
trees_sp = as.SpatialPoints.ppp(trees_ppp)
proj4string(trees_sp) <- CRS('+init=epsg:26978')

sp_dens = smooth_map(trees_sp,
                     breaks = seq(0, 50000, by = 5000),
                     style="fixed", 
                     bandwidth = c(0.01, 0.01))

tm_shape(sp_dens$raster) +
  tm_raster()
```

```{r}
# e) Estimate the K function and plot it along with the envelope for testing CSR.
trees_env = envelope(trees_ppp, Kest, correction = "border")
plot(trees_env)
```

```{r}
# f) Consider only the trees with diameter less than or equal to 20. Repeat part (e)
trees = data.table::fread("homework/trees.csv")
trees_ppp20 = trees[dia <= 20, ppp(x, y, xrange = c(0, 200), yrange = c(0, 200))]
trees_env20 = envelope(trees_ppp20, Kest, correction = "border")
plot(trees_env20)
```

