---
title: "BST 6200 Spatial Statistics and Disease Mapping: Homework 1"
subtitle: "Spring 2020"
author: "Miao Cai <miao.cai@slu.edu>"
date: "1/27/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: readable
    highlight: textmate
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
  .col2 {
    columns: 2 300px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 300px; /* chrome, safari */
    -moz-columns: 2 300px;    /* firefox */
  }
</style>


Data
====
Data description
----------------
- **Outcome variable**: Birth rate for all mother ages (white and black) in 2018.
- **Data source**: [Georgia Department of Public Health's Data Warehouse](https://oasis.state.ga.us/oasis/webquery/qryBirth.aspx).


Read data
---------

```{r message=F}
pacman::p_load(sf, viridisLite, readxl, dplyr, highcharter)

d_white = read_excel("data/dph_wq_Birth_Rate_by_Residence__White_1_27_20_05_21_45_601531_PM.xlsx") %>% 
  slice(3:161) %>% 
  `colnames<-`(c("County", "Birth_rate_white")) %>% 
  mutate(Birth_rate_white = as.numeric(Birth_rate_white))
  
d_black = read_excel("data/dph_wq_Birth_Rate_by_Residence__Black_or_African-American_1_27_20_05_21_58_397429_PM.xlsx") %>% 
  slice(3:161) %>% 
  `colnames<-`(c("County", "Birth_rate_black")) %>% 
  mutate(Birth_rate_black = as.numeric(Birth_rate_black))
```



```{r eval=FALSE,echo=FALSE}
ga = read_sf("data/Georgia.shp") %>% 
  left_join(d_white, by = c("Name" = "County")) %>% 
  left_join(d_black, by = c("Name" = "County"))
ga = left_join(d_white, d_black, by = "County") 
```


Choropleth maps
===============
Introducing `highchartr`
------------------------
The following two choropleth maps are created using the R package [`highchartr`](https://cran.r-project.org/web/packages/highcharter/), which is a wrapper for the [Highcharts](https://www.highcharts.com/) Javascript library. This package can create beautiful choropleth map with interactivity. Move your mouse over each county and you can get the information.

Define color scheme
-------------------
```{r eval=F,echo=F}
discrete_color = data.frame(
  from = seq( 0,  90, 10),
  to   = seq(10, 100, 10),
  color = substring(viridis(10, option = "C", direction = 1), 0, 7)) %>% 
  list_parse()

# Some notes:
# dataClasses = discrete_color
# dataClasses = color_classes(c(seq(0, 10, by = 2), 50))
# dataLabels = list(enabled = TRUE, format = '{point.name}'),
```

```{r}
continuous_color = data.frame(
  q = 0:4/5,
  c = substring(viridis(5, option = "C", direction = 1), 0, 7),
  stringsAsFactors = FALSE) %>% 
  list_parse2()
```

Code to choropleth map in `highchartr`
--------------------------------------
```{r}
p1 = hcmap("countries/us/us-ga-all", 
      data = d_white, 
      value = "Birth_rate_white",
      joinBy = c("name", "County"),
      name = "White birth rate",
      borderColor = "#FAFAFA", 
      borderWidth = 0.5,
      tooltip = list(valueDecimals = 2, 
                     valueSuffix = "%")) %>% 
  hc_colorAxis(stops = continuous_color) %>% 
  hc_legend(layout = "vertical", 
            align = "right", 
            verticalAlign = "top",
            y = 60,
            floating = TRUE, 
            valueDecimals = 0, 
            valueSuffix = "%") %>% 
  #hc_mapNavigation(enabled = TRUE) %>% 
  hc_title(text = 
    "Birth rates among white mothers in Georgia, USA 2018") %>% 
  hc_subtitle(text = 
    "Georgia Department of Public Health's Data Warehouse") %>% 
  hc_add_theme(hc_theme_ft())
```

```{r}
p2 = hcmap("countries/us/us-ga-all", 
      data = d_black, 
      value = "Birth_rate_black",
      joinBy = c("name", "County"),
      name = "Black birth rate",
      borderColor = "#FAFAFA", 
      borderWidth = 0.5,
      tooltip = list(valueDecimals = 2, 
                     valueSuffix = "%")) %>% 
  hc_colorAxis(stops = continuous_color) %>% 
  hc_legend(layout = "vertical", 
            align = "right", 
            verticalAlign = "top",
            y = 60,
            floating = TRUE, 
            valueDecimals = 0, 
            valueSuffix = "%") %>% 
  hc_title(text = 
    "Birth rates among black mothers in Georgia, USA 2018") %>% 
  hc_subtitle(text = 
     "Georgia Department of Public Health's Data Warehouse") %>% 
  hc_add_theme(hc_theme_ft())
```

Demonstration of the two choropleth maps
----------------------------------------
<div class="col2">
```{r echo=FALSE}
p1
```


```{r echo=FALSE}
p2
```
</div>

You could mouse over to see the name of the counties and the corresponding birth rates.


Conclusion
==========
Based on the above two choropleth maps, the birth rates among white and black mothers in Georgia are similar. However, there are minor differences: black mothers tend to have higher birth rates in the south while lower birth rates in the north, although the differences are not substantial.
